\documentclass[11pt, oneside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
%\geometry{landscape}                		% Activate for rotated page geometry
\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex
								
%\usepackage{{Gratzer-Color-Scheme}	% Active to color theorems red and lemmas blue
	
\usepackage{amssymb}

%SetFonts

%SetFonts


\title{Exercise, EEG Oscillations, and Emotional Intelligence in Schizophrenia}
\author{The Author}
%\date{}							% Activate to display a given date or no date

\begin{document}
\maketitle
%\section{}
%\subsection{}

\begin{abstract}
This is where the abstract goes.
\end{abstract}

\section{Introduction}

Exercise and SSD

Exercise and EEG

Oscillations and SSD

EI imbalance and SSD

Gap in Lit

Hypotheses


\section{Methods}

\subsection{Participants} 

\subsection{Design and Procedure}

\subsection{Materials and Measures}

\subsection{EEG data acquisition (Nina) and processing (Berkhan & Genc)}

For EEG data acquisition, BrainVision Recorder was used. We recorded the data in a sound- and electrically- shielded Faraday cage that minimized external noise, using a 32-channel ActiCAP Ag/AgCl active electrode system as well as a BrainAmp amplifier (Brain Products GmbH, Germany). During the recording, we maintained all electrode impedances below 25 kΩ. 31 of the 32 channels were used for scalp EEG recording, while the Oz electrode was placed on the left lower back and used as a drop-down ECG electrode. We arranged the electrodes according to the International 10–20 electrode system (see Figure XX), with the FCz electrode as reference and AFz as ground, and recorded with a sampling rate of 500 Hz. Participants were instructed to relax, stay awake and minimize movement during all recordings. In particular, they were instructed to keep their head still, avoid grinding their teeth and squinting. We furthermore placed a Respiration Belt (Brain Products GmbH, Germany) around participants’ chests to track respiration during the whole experiment. 

\subsection{Statistical analysis}


\section{Results}

\section{Discussion}

\bibliographystyle{plainnat}
\bibliography{refs} % refs.bib file with references


\end{document}  